\section{Introduction}

The goal of this first assignment was to implement an Iterative Improvement
algorithm and a Variable Neighbourhood Descent algorithm. Multiple rules were
implemented for each algorithm in order to test their efficiency. For each
matrix instance we were given, we also have the best-known score we should try
to approach. For each combination of rules, we also had to run the 2 programs
on each instance. The results of each combination were used to compute
statistical tests with R.

\subsection{Hierarchy}

\noindent The file tree below contains a short description of the assignment's
directories:

\begin{framed}
\dirtree{%
.1 Assignment1.
.2 best\_known - best-known scores for each instance.
.2 experiments - contains the results of each experiment executed on each instance.
.2 instances - matrix instances.
.2 report - report's LaTeX sources.
.2 src.
    .3 IterativeImprovement - classes related to the first exercise.
    .3 rules - classes related to rules (initialisation, pivot, neighbourhood).
    .3 tests - tests to verify that implemented features are working individually.
    .3 VariableNeighbourhoodDescent - classes related to the second exercise.
.2 subject.
.2 CMakeLists.txt - Used by cmake in order to generate a Makefile.
}
\end{framed}

\subsection{Compilation}

\noindent In order to compile the programs, cmake is necessary and the
instructions below must be followed.

\begin{lstlisting}[language=bash]
> mkdir out
> cd out
> cmake ..
> make lop vnd
\end{lstlisting}

\newpage
